Honey Bee Colonies Impacted by Varroa, American Fouldbrood and Global Warming

### 2024-06-24 docs/ManuscriptGrp10.qmd
Docs Folder  Cleaned up: All R code and Data Sets tested and verified 06.24.24

```{r}
library(readxl)
library(tidyverse)
library(dplyr)

```{r}
# Output cleaned data file to a csv file.
hcny_data <- read.csv("C:/Users/ecruz/OneDrive/Documents/UTSA - Data Science Program/Semester Classes/Practicum II Repository/P2-Practicum-II-Portfolio-EdwardCruz/docs/hcny_CleanDraft.csv", header = FALSE, stringsAsFactors = FALSE)

#view(hcny_data)
```

```{r}
# Read the CSV file
hcny_data <- read.csv("C:/Users/ecruz/OneDrive/Documents/UTSA - Data Science Program/Semester Classes/Practicum II Repository/P2-Practicum-II-Portfolio-EdwardCruz/docs/hcny_CleanDraft.csv", header = FALSE, stringsAsFactors = FALSE)

```{r}
# Merge the first two rows to create a proper header
header <- hcny_data[1:2, ]
header <- sapply(header, function(x) paste(na.omit(x), collapse = " "))
colnames(hcny_data) <- header

# Remove the first two rows (header rows) and any completely blank rows
data_clean2 <- hcny_data[-c(1, 2), ]

# Remove rows that are completely blank
data_clean2 <- data_clean2[rowSums(is.na(data_clean2) | data_clean2 == "") != ncol(data_clean2), ]

# Convert columns to numeric where applicable (if they contain percentage values)
hcnydata_cleaned <- data_clean2 %>%
  mutate(across(.cols = -1, .fns = ~ as.numeric(gsub("[^0-9.-]", "", .))))

# Assuming hcny_data_cleaned is your cleaned data frame
colnames(hcnydata_cleaned) <- c("state", "varroa_mites", "other_pests", "disease", "pesticides", "other", "unknown")
```

```{r}
# Write the cleaned data to a new CSV file
write.csv(hcnydata_cleaned, "C:/Users/ecruz/OneDrive/Documents/UTSA - Data Science Program/Semester Classes/Practicum II Repository/P2-Practicum-II-Portfolio-EdwardCruz/docs/hcnydata_cleaned.csv", row.names = FALSE)

# View the updated data frame
view(hcnydata_cleaned)
```

```{r}
ggplot(hcnydata_cleaned, aes(varroa_mites)) + geom_histogram()
```

>>>>

---
title: "loading/cleaning/wrangling/exploring"
author: "Edward Cruz"
format: docx
---

```{r}
library(readxl)
update.packages("tidyverse")
library(tidyverse)
update.packages("dplyr")
library(dplyr)
update.packages('mgcv')
library(mgcv)
```

```{r}
# Output cleaned data file to a csv file.
hcny_data <- read.csv("C:/Users/ecruz/OneDrive/Documents/UTSA - Data Science Program/Semester Classes/Practicum II Repository/P2-Practicum-II-Portfolio-EdwardCruz/docs/hcny_CleanDraft.csv", header = FALSE, stringsAsFactors = FALSE)
```

```{r}
head(hcny_data)
```

```{r}
str(hcny_data)
```

```{r}
summary(hcny_data)
```

```{r}
view(hcny_data)
```

```{r}
# Merge the first two rows to create a proper header
header <- hcny_data[1:4, ]
header <- sapply(header, function(x) paste(na.omit(x), collapse = " "))
colnames(hcny_data) <- header
head(header)
```

```{r}
# Remove the first two rows (header rows) and any completely blank rows
hcnydata_cleaned <- hcny_data[-c(1,2), ]
head(hcnydata_cleaned)
```

```{r}
# Assuming hcny_data_cleaned is your cleaned data frame
colnames(hcnydata_cleaned) <- c("State", "Varroa_Mites", "Other_Parasites_1/", "Disease_2/", "Pesticides", "Other_3/", "Unknown")
```

```{r}
head(hcnydata_cleaned)
```

```{r}
# Convert specified columns to numeric
hcnydata_cleaned <- hcnydata_cleaned %>%
  mutate(across(c("Varroa_Mites", "Other_Parasites_1/", "Disease_2/", "Pesticides", "Other_3/", "Unknown"), as.numeric))
```

```{r}
# View the updated data frame
hcnydata_cleaned<-hcnydata_cleaned[-1,]
```

```{r}
head(hcnydata_cleaned)
```

```{r}
# Verify the changes
str(hcnydata_cleaned)
```

```{r}
head(hcnydata_cleaned)
```

```{r}
view(hcnydata_cleaned)
```

```{r}
ggplot(hcnydata_cleaned, aes(Varroa_Mites)) + geom_histogram()
```

```{r}
ggplot(hcnydata_cleaned, aes(x = Varroa_Mites)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  labs(title = "Histogram of Varroa Mites",
       x = "Varroa Mites",
       y = "Frequency")
```

```{r}
# Example scatter plot of two variables
ggplot(hcnydata_cleaned, aes(x = Varroa_Mites, y = "Desease_2/")) +
  geom_point() +
  labs(title = "Scatter Plot of Varroa Mites vs Disease",
       x = "Varroa Mites",
       y = "2 American and European Foul Brood")
```

```{r}
view(hcnydata_cleaned)
```

```{r}
# Write the cleaned data to a new CSV file
write.csv(hcnydata_cleaned, "C:/Users/ecruz/OneDrive/Documents/UTSA - Data Science Program/Semester Classes/Practicum II Repository/P2-Practicum-II-Portfolio-EdwardCruz/docs/hcnydata_cleaned.csv", row.names = FALSE)
```

\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*Conclusion- all code below is test code right now\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*
